---
title: "Hayman Fire Recovery"
author: "AJ Lee"
date: "Feb/2/2022"
output: html_document


---

```{r setup, warning=F,message=F}
library(tidyverse)
library(tidyr)
library(ggthemes)
library(lubridate)

# Now that we have learned how to munge (manipulate) data
# and plot it, we will work on using these skills in new ways

knitr::opts_knit$set(root.dir='../docs')

```


```{r dataread, warning=F,message=F}
####-----Reading in Data and Stacking it ----- ####
#Reading in files
files <- list.files('../data',full.names=T)


#Read in individual data files
ndmi <- read.csv(files[1]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndmi')


ndsi <- read.csv(files[2]) %>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndsi')

ndvi <- read.csv(files[3])%>% 
  rename(burned=2,unburned=3) %>%
  mutate(data='ndvi')

# Stack as a tidy dataset
full_long <- rbind(ndvi,ndmi,ndsi) %>%
  gather(key='site',value='value',-DateTime,-data) %>%
  filter(!is.na(value))


```


## Question 1 

1) What is the correlation between NDVI and NDMI? - here I want you to
convert the full_long dataset in to a wide dataset using the 
function "spread" and then make a plot that shows the correlation as a
function of if the site was burned or not (x axis should be ndmi)
You should exclude winter months and focus on summer months.  
**A: NDVI and NDMI have a positive correlation due to moisture is an important
     factor for vegetation. In particular, unburned area showed a clearer relation
     than burned area.**

```{r, warning=F,message=F}

full_wide <- spread(data=full_long, key='data', value='value') %>%
  filter_if(is.numeric, all_vars(!is.na(.))) %>%
  mutate(month = month(DateTime),
         year = year(DateTime))

summer_only <- filter(full_wide, month %in% c(6,7,8,9))

ggplot(summer_only,aes(x=ndmi, y=ndvi, color=site)) + 
  geom_point() + 
  theme_few() + 
  scale_color_few() + 
  theme(legend.position=c(0.8,0.8))

```



## Question 2 

2) What is the correlation between average NDSI (normalized
 snow index) for January - April and average NDVI for June-August?
In other words, does the previous year's snow cover influence vegetation
 growth for the following summer?  
 **A: NDVI and NDSI do not show a significant correlation in this research, but
   a difference in the correlation between burned and un-burned area can be seen** 
 
```{r, warning=F, message=F}

NDSI_NDVI <- mutate(full_long, month=month(DateTime), 
                    year=year(DateTime)) %>%
            filter((data=='ndsi' & month %in% c(1:4))|(data=='ndvi' & month %in% c(6,7,8))) %>% 
            group_by(data,site,year) %>% 
            summarize(mean_value=mean(value)) %>% 
            spread(key='data', value='mean_value')

ggplot(NDSI_NDVI, aes(x=ndsi, y=ndvi, color=year)) +
  geom_point() +
  theme_few() +
  theme(legend.position=c(0.8,0.2)) +
  facet_wrap(~site)

```


## Question 3

3) How is the snow effect from question 2 different between pre- and post-burn
   and burned and unburned?  
   **A: As can be seen from the graph of burned area above, the effect of snow on
   vegetation appears to have decreased sharply after the fire, while the effect
   is still there over time in unburned areas.**


## Question 4

4) What month is the greenest month on average? Does this change in the burned 
   plots after the fire?  
**A: Before the wildfire, September was the greenest month on average, and after 
   the fire, August was the greenest month in burned area.**

```{r, warning=F,message=F}

Green_ndvi <- gather(ndvi, key='site',value='NDVI',-DateTime,-data) %>%
              filter(!is.na(NDVI)) %>% 
              mutate(month = month(DateTime)) %>%
              group_by(site, month) %>% 
              summarize(mean_NDVI=mean(NDVI))

ggplot(Green_ndvi, aes(x=month, y=mean_NDVI, color=site)) +
  geom_point() +
  geom_line(linetype=2) +
  theme_few() +
  scale_color_few() +
  theme(legend.position=c(0.5,0.2))

```



## Question 5) 

5) What month is the snowiest on average?  
**A: Before the wildfire, January was the snowiest month on average, and after the fire, February was the snowiest month on average in burned area.**
```{r, warning=F,message=F}

Snow_ndsi <- gather(ndsi, key='site', value='NDSI', -DateTime, -data) %>% 
             filter(!is.na(NDSI)) %>% 
             mutate(month = month(DateTime)) %>% 
             group_by(site, month) %>% 
             summarize(mean_NDSI=mean(NDSI))

ggplot(Snow_ndsi, aes(x=month, y=mean_NDSI, color=site)) +
  geom_point() +
  geom_line(linetype=2) +
  theme_few() +
  scale_color_few() +
  theme(legend.position=c(0.5,0.8))

```


## Bonus Question: Redo all problems with `spread` and `gather` using modern tidyverse syntax. 


## Bonus Question: Use Climage Engine to pull the same data for the assignment, but updated with 2020/2021 data.




